Sparse Word Embedding for Multi-domain Machine Translation

### 2019-04-24 2019/emnlp/sparse.tex
following our discussion

\section{Introduction}
\fyTodo{Merge refs and order chronologically}
\fyTodo{Homogeneize cite keys, avoid arxiv}

@@

In \fyTodostar{Specify DA}{Multi-domain Machine Translation}, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpora EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB. " (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}). Using context information is not sufficient to differ these polysemies. Therefore using the same word embedding for polysemies obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
\begin{table}[h]
\resizebox{\columnwidth}{!}{
\begin{tabular}{|l|l|l|}
\hline
English Words & French translation & Domain \\ \hline
security & titre financier & Financial \\
security & s\'ecuti\'e & IT\\
\hline
joint & articulaire & medical\\
joint & conjoint & banking\\
\hline
chair & chaise & general \\
chair & pr\'esider, pr\'esident & banking \\
\hline
\end{tabular}
}
\caption{Polysemies}
\label{table:polysemy}
\end{table}

\begin{table}[h]
\resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||}
\hline
English sentence & French translation & Corpora \\ 
\hline
"Flu-like" symptoms such as headaches, \colorbox{red}{joint} pains, feelings of weakness, dizziness, and tiredness may occur, especially at the start of treatment. & Des symptômes grippaux tels que céphalées, douleurs \colorbox{red}{articulaires}, sensation de faiblesse, vertige et asthénie, peuvent survenir, en particulier en début de traitement. & EMEA \\
\hline
In my opinion, Mr President, this \colorbox{red}{joint} debate fully bears this out. & Cette discussion \colorbox{red}{commune} en est, Monsieur le Président, la preuve vivante. & EPPS \\
\hline
The increasing number of central banks using the same IT platform for their own purposes has provided unique opportunities to undertake \colorbox{red}{joint} projects , exploit synergies and better manage the relationship with the vendor. & Le nombre croissant de banques centrales utilisant la même plate-forme informatique pour leurs besoins propres a offert des occasions uniques de lancer des projets \colorbox{red}{communs}, d'exploiter des synergies et de mieux gérer les relations avec le fournisseur. & ECB \\
\hline
\end{tabularx}
}
\caption{Translation}
\label{table:polysemy}
\end{table}
\fyTodo{Typical plan: the problems, existing solutions and their limits, our idea and contributions}
In this work, we present a new design for word embedding that is composed of different regions which are activate and deactivate corresponding to domains. 
\\ To evaluate, we use 3 corpora of 3 different domains: Medical, Banking and Administration and 2 different architectures: Transformer and Attentional recurrent network. The test sets are selectioned from shared task each domains and two other general test sets such as news commentary and IWSLT. The reader could find more details in Experimetal section \ref{subsec:data}
\fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}

\section{Related Work \label{sec:related_work}}
\fyTodo{Add standard labels to sections}
\fyTodo{Related work goes last}
\fyTodo{Compare also to Peng 2017}

@@

Figure~\ref{fig:network} presents our proposed design. During backprogation, the \Todo{rotation}{Why ?} matrix receives vanishing gradient at regions corresponding to deactivated regions in the word embedding. Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they do not correspond.

@@

\subsection{NMT model configuration}
We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{NIPS2017_7181}.
More precisely, for the Transformer we use the same settings as described in \cite{NIPS2017_7181}. For the recurrent architecture, in order to fairly compare to the work of \citet{D18-1041}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512. The implementation can be found in OpenNMT-tf \cite{P17-4012}.
\fyTodo{Language pairs, BPE number, optimizer}
\fyTodo{Motivate the split}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region. If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

@@

We also take in account the problem of overfitting that could occur for small corpora. Indeed, real data are not always balanced for every domain. For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English-German. If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting. Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora. \fyTodo{Refs on this ?}

\subsection{Results}
\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
 \hline
 ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
 \hline
 Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BPE-detokenized BLEU score in language pair English - French. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by pretrained domain classififer}
\label{tab:1}
\end{figure}

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & 18.85 \\
 \hline
 EMEA-finetuned & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse(1) & 70.55 & 26.17 & 68.07 & 21.50 & 23.30 & 14.05 & 16.64 & 19.04\\
 \hline
 Sparse(2) & 70.67 & 26.24 & 66.97 & & & 14.05 & 16.64 & 19.04 \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BLEU score for Transformer in language pair English - German. The notation is same as figure \ref{tab:1}}
\label{tab:2}
\end{figure}

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse & & & & & & & & \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
 \hline 
\end{tabularx}}
\end{center}
\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French. Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LSTM}
\label{tab:3}
\end{figure}

\begin{figure}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & \\
 \hline
 Sparse & & & & & & & \\
 \hline
 Domain Control & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & & & & & & & \\
 \hline 
\end{tabularx}}
\end{center}
\caption{BLEU score for Recurrent network on language pair English - German. Notations are the same as for figure~\ref{tab:3}}
\label{tab:4}
\end{figure}
\subsection{Word embedding analysis}

>>>>

\section{Introduction \label{sec:introduction}}
\fyDone{Merge refs and order chronologically}
\fyTodo{Homogeneize cite keys, avoid arxiv}

@@

In \fyTodostar{Specify DA}{Multi-domain Machine Translation}, ones have to deal with polysemies whose meaning can be totally different in different domains, for example, the word "chair" means an household tool in "if you are taking a medicine which may cause low blood pressure when rising from a chair or bed " (extracted from medical corpus EMEA \cite{Tiedemann2009RANLP5}) while the word "chair" means to preside over a meeting in following sentence "The President of the ECB or , in his absence , the Vice-President , chairs the meetings of the Governing Council , the Executive Board and the General Council of the ECB. " (extracted from European Central Bank corpora \cite{Tiedemann2009RANLP5}). Using context information is not sufficient to differ these polysemies. Therefore using the same word embedding for polysemies obstructs the model from differentiating the meaning of a polysemy in one domain from it's other meaning in the other domain.
\begin{table}[h]
\resizebox{\columnwidth}{!}{
\begin{tabular}{|l|l|l|}
\hline
English Words & French translation & Domain \\ \hline
security & titre financier & Financial \\
security & s\'ecuti\'e & IT\\
\hline
joint & articulaire & medical\\
joint & conjoint & banking\\
\hline
chair & chaise & general \\
chair & pr\'esider, pr\'esident & banking \\
\hline
\end{tabular}
}
\caption{Polysemies}
\label{table:polysemy}
\end{table}

\begin{table}[h]
\resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||}
\hline
English sentence & French translation & Corpus \\ 
\hline
"Flu-like" symptoms such as headaches, \colorbox{red}{joint} pains, feelings of weakness, dizziness, and tiredness may occur, especially at the start of treatment. & Des symptômes grippaux tels que céphalées, douleurs \colorbox{red}{articulaires}, sensation de faiblesse, vertige et asthénie, peuvent survenir, en particulier en début de traitement. & EMEA \\
\hline
In my opinion, Mr President, this \colorbox{red}{joint} debate fully bears this out. & Cette discussion \colorbox{red}{commune} en est, Monsieur le Président, la preuve vivante. & EPPS \\
\hline
The increasing number of central banks using the same IT platform for their own purposes has provided unique opportunities to undertake \colorbox{red}{joint} projects , exploit synergies and better manage the relationship with the vendor. & Le nombre croissant de banques centrales utilisant la même plate-forme informatique pour leurs besoins propres a offert des occasions uniques de lancer des projets \colorbox{red}{communs}, d'exploiter des synergies et de mieux gérer les relations avec le fournisseur. & ECB \\
\hline
\end{tabularx}
}
\caption{Translation}
\label{table:polysemy}
\end{table}
\fyTodo{Typical plan: the problems, existing solutions and their limits, our idea and contributions}
In this work, we present a new design for word embedding that is composed of different regions which are activate and deactivate corresponding to domains. 
\\ To evaluate, we use 3 corpora of 3 different domains: Medical, Banking and Administration and 2 different architectures: Transformer and Attentional recurrent network. The test sets are selectioned from shared task each domains and two other general test sets such as news commentary and IWSLT. The reader could find more details in Experimetal section \ref{subsec:data}
\fyTodo{More precise: two architectures, two languages pairs, multiple domains, also improves generic}

@@

Figure~\ref{fig:network} presents our proposed design. During backprogation, the \fyTodo{fix notation}{projection matrix $P$} receives vanishing gradient at regions corresponding to deactivated regions in the word embedding. Those regions are also vanished in forward step, thus do not interfere the training on the domain to which they do not correspond.

@@

\subsection{NMT model configuration}
We evaluate our method on two state-of-the-art encoder-decoder architectures: a recurrent architecture based on attentional birectional Long Short-Term Memory (LSTM) cells and the Transformer architecture of \citet{NIPS2017_7181}.
More precisely, for the Transformer we use the same settings as described in \cite{NIPS2017_7181}. For the recurrent architecture, in order to fairly compare to the work of \citet{D18-1041}, we use a standard configuration with one layer bidirectional LSTM with a encoder hidden size of 1024; one attentional unidirectional LSTM layer with same hidden size in decoder; and word embeddings of size 512. The implementation can be found in OpenNMT-tf \cite{P17-4012}.
\fyTodo{Language pairs, BPE number, optimizer}
\fyTodo{Motivate the split - discuss experimentally embedding size}
For the implementation of sparse embeddings, we split the vector embedding of size 512 into four regions $[8,8,8,488]$, corresponding respectively to EMEA, EPPS, ECB domain and generic region. If sentence comes from domain $i$, regions for domains $j,\forall j \neq i$ will be masked as 0. We then use a dense layer of size 512 to fuse the active domain region and the generic region.

@@

We also take in account the problem of overfitting that could occur for small corpora. Indeed, real data are not always balanced for every domain. For instance, in our experiments, the size of corpora are very different where EPPS is largest with approximately 2 millions sentences, EMEA contains about 1 million sentences and ECB is the smallest with only 200 000 sentences for English French and 100 000 for English-German. If we train each corpora with same number of batches, ECB will be trained too oftenly, that easily leads to overfitting. Therefore, we perform backpropagation on corpora with frequencies according to sizes of corpora. \fyTodo{Refs on this ? or contrast ?}

\subsection{Results}
\begin{table}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 67.69 & 37.5 & 53.49 & 34.86 & 30.96 & 20.98 &  & 25.7 \\
 \hline
 EMEA-finetuned & 76.77 & 17.16 & 11.99 & 29.58 & 14.25 & 9.92 &  & 11.10 \\
 \hline
 EPPS-finetuned & 20.86 & 37.04 & 24.53 & 31.86 & 30.97 & 20.72 &  & 11.1 \\
 \hline
 ECB-finetuned & 9.2 & 18.39 & 55.45 & 12.13 & 16.12 & 8.71 &  & 7.19 \\
 \hline
 Sparse(1) & 70.45 & 38.23 & 54.69 & 34.66 & 31.24 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(2) & 69.91 & 38.05 & 54.2 & 35.22 & 31.25 & 21.36 & 28.12 & 26.02 \\
 \hline
 Sparse(3) &  &  &  &  &  &  &  & \\
 \hline
 Domain Control & 67.87 & 37.31 & 54.14 & 33.63 & 31.09 &  & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BPE-detokenized BLEU score for English-French with the Transformer architecture. Sparse(1) uses correponding domain tag for EMEA, EPPS, ECB test sets and generic tag for newstest and IWSLT. Sparse(2) uses generic tag for all of test sets. Sparse(3) uses tag predicted by a pre-trained domain classififer}
\label{tab:1}
\end{table}

\begin{table}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & 64.57 & 26.47 & 68.67 & 21.77 & 23.50 & 13.9 & 17.04 & 18.85 \\
 \hline
 EMEA-finetuned & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse(1) & 70.55 & 26.17 & 68.07 & 21.50 & 23.30 & 14.05 & 16.64 & 19.04\\
 \hline
 Sparse(2) & 70.67 & 26.24 & 66.97 & & & 14.05 & 16.64 & 19.04 \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
\end{tabularx}}
\end{center}
\caption{BLEU score for Transformer in language pair English - German. Notations are same as Table~\ref{tab:1}}
\label{tab:2}
\end{table}

\begin{table}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & newstest 2014 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & & \\
 \hline
 Sparse & & & & & & & & \\
 \hline
 Domain Control & & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & 68.76 & 35.71 & 52.75 & 32.03 & 29.81 & 19.58 & 24.63 & 23.28 \\
 \hline 
\end{tabularx}}
\end{center}
\caption{bpe-detokenized BLEU score for Recurrent network architecture on language pair English-French. Notations as same as figure \ref{tab:1} except that other models from WDCMT of \cite{D18-1041} use Bidirectional-LSTM}
\label{tab:3}
\end{table}

\begin{table}[H]
\begin{center}
 \resizebox{\columnwidth}{!}{\begin{tabularx}{\textwidth}{|| X | X | X | X | X | X | X | X ||} 
 \hline
 Models & EMEA test & EPPS test & ECB test & khresmoi (EMEA domain) & test 2007 (EPPS domain) & newstest 2009 & IWSLT test 2010 \\ [0.5ex] 
 \hline\hline
 Mixed & & & & & & & \\
 \hline
 EMEA-finetuned & & & & & & & \\
 \hline
 EPPS-finetuned & & & & & & & \\
 \hline
 ECB-finetuned & & & & & & & \\
 \hline
 Sparse & & & & & & & \\
 \hline
 Domain Control & & & & & & & \\
 \hline
 WDCMT (conditional GRU) & & & & & & & \\
 \hline 
\end{tabularx}}
\end{center}
\caption{BLEU scores for RNN architecture on English-German. Notations are the same as for figure~\ref{tab:3}}
\label{tab:4}
\end{table}

\subsection{Word embedding analysis}

\section{Related Work \label{sec:related_work}}
\fyDone{Add standard labels to sections}
\fyDone{Related work goes last}
\fyTodo{Compare also to Peng 2017}
